---
title: "Binary package production for GCP/AnVIL"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Binary package production for GCP/AnVIL"}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---
```{r setup,echo=FALSE,results="hide"}
suppressMessages({
suppressPackageStartupMessages({
library(BiocStyle)
library(BiocBuildTools)
library(BiocPkgTools)
library(igraph)
library(dplyr)
library(magrittr)
})
})
```

# Basic concepts

When R is used within a Docker container, users may confront
the need to install additional packages.  This can be time-consuming
when compilation is required.  To solve this, we have created
a repository of container-based binary packages.  Installing
these during R sessions is fast and only involves file transfer.

# A container for use with NHGRI AnVIL

We will use the workspace

```
https://app.terra.bio/#workspaces/landmarkanvil2/biocbinaries_apr2020
```

to run a custom environment with 64 CPU, 240 GB RAM, and 300 GB disk.

We use the runtime container 

```
us.gcr.io/anvil-gcr-public/anvil-rstudio-bioconductor:0.0.4
```

# The manifest for Bioconductor 3.10 software packages

We use the following function to generate a character
vector with names of all software packages in a given
release.
```{r getman}
get_bioc_packagelist = function(rel = "RELEASE_3_10") {
 system("git clone git@git.bioconductor.org:admin/manifest")
 owd = getwd()
 setwd("manifest")
 on.exit(setwd(owd))
 system(paste("git checkout ", rel))
 proc_software.txt = function() {
  x = readLines("software.txt")[-1]  # first line is comment
  nn = which(nchar(x)==0)
  tmp = x[-nn]
  gsub("Package: ", "", tmp)
 }
 proc_software.txt()
}
```
We'll generate the vector of package names and copy it
to a google bucket for retrieval in our AnVIL session.

```{r gepks,eval=FALSE}
software_3.10_2020_04_28 = get_bioc_packagelist()
save(software_3.10_2020_04_28, file="software_3.10_2020_04_28.rda")
system("gsutil cp software_3.10_2020_04_28.rda gs://biocbbs_2020a")
```

We will use this vector as input to `BiocManager::install` with
Ncpus = 45.

# Tasks within AnVIL

We set up the runtime based on
```
us.gcr.io/anvil-gcr-public/anvil-rstudio-bioconductor:0.0.4
```

In Rstudio, we retrieve the vector of package names.

```
R version 3.6.1 (2019-07-05) -- "Action of the Toes"
Copyright (C) 2019 The R Foundation for Statistical Computing
Platform: x86_64-pc-linux-gnu (64-bit)

R is free software and comes with ABSOLUTELY NO WARRANTY.
You are welcome to redistribute it under certain conditions.
Type 'license()' or 'licence()' for distribution details.

R is a collaborative project with many contributors.
Type 'contributors()' for more information and
'citation()' on how to cite R or R packages in publications.

Type 'demo()' for some demos, 'help()' for on-line help, or
'help.start()' for an HTML browser interface to help.
Type 'q()' to quit R.

> getwd()
[1] "/home/rstudio"
> system("gsutil cp gs://biocbbs_2020a/software_3.10_2020_04_28.rda .")
Copying gs://biocbbs_2020a/software_3.10_2020_04_28.rda...
/ [1 files][ 11.6 KiB/ 11.6 KiB]                                                
Operation completed over 1 objects/11.6 KiB.                                     
> dir()
[1] "entrypoint.out"               "kitematic"                    "software_3.10_2020_04_28.rda" "welder.log"                  
> load("software_3.10_2020_04_28.rda")
> length(software_3.10_2020_04_28)
[1] 1823
```



